### 2017-06-09 affiliations.tex

$^{1}$INRIA Bordeaux Sud-Ouest
      Talence, France --
      Institut des Maladies Neurodégénératives, Université de Bordeaux,
      CNRS UMR 5293, Bordeaux, France -- 
      LaBRI, Université de Bordeaux, Institut Polytechnique de Bordeaux,
      CNRS UMR 5800, Talence, France
      %%
      •
      %%
$^{2}$Centre de Biophysique Moléculaire,
      CNRS UPR4301, Orléans --
      Synchrotron SOLEIL, Division Expériences,
      Gif sur Yvette, France
      %%
\par