### 2017-06-16 authors.tex
Updated authors list

\textbf{Nicolas P. Rougier}$^{1\dagger\S\star}$\href{http://orcid.org/0000-0002-6972-589X}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Konrad Hinsen}$^{2\dagger}$\href{http://orcid.org/0000-0003-0330-9428}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Frédéric Alexandre}$^{3\S}$\href{http://orcid.org/0000-0002-6113-1878}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Thomas Arildsen}$^{4\dagger}$\href{http://orcid.org/0000-0003-3254-3790}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Lorena Barba}$^{5\dagger}$\href{http://orcid.org/0000-0001-5812-2711}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Fabien C. Y. Benureau}$^{6\ddagger}$\href{http://orcid.org/0000-0003-4083-4512}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{C. Titus Brown}$^{7\dagger}$\href{http://orcid.org/0000-0001-6001-2677}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Pierre de Buyl}$^{8\dagger}$\href{http://orcid.org/0000-0002-6640-6463}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Ozan Çağlayan}$^{9\ddagger}$\href{http://orcid.org/0000-0002-5992-3470}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Andrew Davison}$^{10\ddagger}$\href{http://orcid.org/0000-0002-4793-7541}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Marc André Delsuc}$^{11\ddagger}$\href{http://orcid.org/0000-0002-1400-5326}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Georgios Detorakis}$^{12\ddagger\S}$\href{http://orcid.org/0000-0001-5891-1702}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Damien Drix}$^{13\ddagger}$\href{http://orcid.org/0000-0003-4107-5693}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Pierre Enel}$^{14\ddagger}$\href{http://orcid.org/0000-0001-8983-6223}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Benoît Girard}$^{15\ddagger\S}$\href{http://orcid.org/0000-0002-8117-7064}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Emmanuelle Gouillart}$^{16\dagger}$\href{http://orcid.org/0000-0002-3869-9967}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Olivier Grisel}$^{17\ddagger}$\href{http://orcid.org/0000-0002-8226-4012}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Olivia Guest}$^{18\dagger}$\href{http://orcid.org/0000-0002-1891-0972}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Matt Hall}$^{19\ddagger}$\href{http://orcid.org/0000-0002-9530-5477}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Rafael Neto Henriques}$^{20\S}$,
%%
\textbf{Xavier Hinaut}$^{21\ddagger}$\href{http://orcid.org/0000-0002-1924-1184}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Mehdi Khamassi}$^{22\ddagger\S}$\href{http://orcid.org/0000-0002-2515-1046}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Almar Klein}$^{23\ddagger}$\href{http://orcid.org/0000-0002-9978-2780}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Andrei Maksimov}$^{24\S}$,
%%
\textbf{Pietro Marchesi}$^{25\ddagger}$\href{http://orcid.org/0000-0001-5955-6909}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Dan McGlinn}$^{26\ddagger}$\href{http://orcid.org/0000-0003-2359-3526}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Christoph Metzner}$^{27\S}$\href{http://orcid.org/0000-0002-5933-2101}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Owen Petchey}$^{28\S}$\href{http://orcid.org/0000-0002-7724-1633}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Hans Ekkehard Plesser}$^{29\ddagger}$\href{http://orcid.org/0000-0001-7843-5993}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Timothée Poisot}$^{30\dagger}$\href{http://orcid.org/0000-0002-0735-5184}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Karthik Ram}$^{31\dagger}$\href{http://orcid.org/0000-0002-0233-1757}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Yoav Ram}$^{32\ddagger}$\href{http://orcid.org/0000-0002-9653-4458}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Etienne Roesch}$^{33\ddagger}$\href{http://orcid.org/0000-0002-8913-4173}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Cyrille Rossant}$^{34\ddagger}$\href{http://orcid.org/0000-0003-2069-9093}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Vahid Rostami}$^{35\S}$,
%%
\textbf{Aaron Shifman}$^{36\ddagger\S}$\href{http://orcid.org/0000-0003-2140-7590}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Joseph Stachelek}$^{37\S}$\href{http://orcid.org/0000-0002-5924-2464}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Marcel Stimberg}$^{38\ddagger}$\href{http://orcid.org/0000-0002-2648-4790}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Frank Stollmeier}$^{39\S}$\href{http://orcid.org/0000-0003-4858-0895}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Federico Vaggi}$^{40\ddagger}$\href{http://orcid.org/0000-0001-8100-158X}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Guillaume Viejo}$^{41\S}$,
%%
\textbf{Julien Vitay}$^{42\ddagger\S}$\href{http://orcid.org/0000-0001-5229-2349}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Anya Vostinar}$^{43\ddagger}$\href{http://orcid.org/0000-0001-7216-5283}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Roman Yurchak}$^{44\ddagger}$\href{http://orcid.org/0000-0002-2565-4444}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Tiziano Zito}$^{45\dagger}$\\