### 2017-06-26 authors.tex

\textbf{Nicolas P. Rougier}$^{1\dagger\S\star}$\href{http://orcid.org/0000-0002-6972-589X}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Konrad Hinsen}$^{2\dagger\star}$\href{http://orcid.org/0000-0003-0330-9428}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Frédéric Alexandre}$^{1\S}$\href{http://orcid.org/0000-0002-6113-1878}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Thomas Arildsen}$^{3\dagger}$\href{http://orcid.org/0000-0003-3254-3790}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Lorena Barba}$^{4\dagger}$\href{http://orcid.org/0000-0001-5812-2711}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Fabien C. Y. Benureau}$^{1\ddagger}$\href{http://orcid.org/0000-0003-4083-4512}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{C. Titus Brown}$^{5\dagger}$\href{http://orcid.org/0000-0001-6001-2677}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Pierre de Buyl}$^{6\dagger}$\href{http://orcid.org/0000-0002-6640-6463}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Ozan Caglayan}$^{7\ddagger}$\href{http://orcid.org/0000-0002-5992-3470}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Andrew Davison}$^{8\ddagger}$\href{http://orcid.org/0000-0002-4793-7541}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Marc André Delsuc}$^{9\ddagger}$\href{http://orcid.org/0000-0002-1400-5326}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Georgios Detorakis}$^{10\ddagger\S}$\href{http://orcid.org/0000-0001-5891-1702}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Damien Drix}$^{11\ddagger}$\href{http://orcid.org/0000-0003-4107-5693}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Pierre Enel}$^{12\ddagger}$\href{http://orcid.org/0000-0001-8983-6223}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Benoît Girard}$^{13\ddagger\S}$\href{http://orcid.org/0000-0002-8117-7064}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Olivia Guest}$^{14\dagger}$\href{http://orcid.org/0000-0002-1891-0972}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Matt G. Hall}$^{15\ddagger}$\href{http://orcid.org/0000-0002-9530-5477}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Rafael Neto Henriques}$^{16\S}$,
%%
\textbf{Xavier Hinaut}$^{1\ddagger}$\href{http://orcid.org/0000-0002-1924-1184}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Mehdi Khamassi}$^{13\ddagger\S}$\href{http://orcid.org/0000-0002-2515-1046}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Almar Klein}$^{17\ddagger}$\href{http://orcid.org/0000-0002-9978-2780}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Pietro Marchesi}$^{18\ddagger}$\href{http://orcid.org/0000-0001-5955-6909}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Dan McGlinn}$^{19\ddagger}$\href{http://orcid.org/0000-0003-2359-3526}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Christoph Metzner}$^{20\S}$\href{http://orcid.org/0000-0002-5933-2101}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Owen L. Petchey}$^{21\S}$\href{http://orcid.org/0000-0002-7724-1633}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Hans Ekkehard Plesser}$^{22\ddagger}$\href{http://orcid.org/0000-0001-7843-5993}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Timothée Poisot}$^{23\dagger}$\href{http://orcid.org/0000-0002-0735-5184}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Karthik Ram}$^{24\dagger}$\href{http://orcid.org/0000-0002-0233-1757}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Yoav Ram}$^{25\ddagger}$\href{http://orcid.org/0000-0002-9653-4458}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Etienne Roesch}$^{26\ddagger}$\href{http://orcid.org/0000-0002-8913-4173}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Cyrille Rossant}$^{27\ddagger}$\href{http://orcid.org/0000-0003-2069-9093}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Vahid Rostami}$^{28\S}$\href{http://orcid.org/0000-0002-3851-0220}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Aaron Shifman}$^{29\ddagger\S}$\href{http://orcid.org/0000-0003-2140-7590}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Joseph Stachelek}$^{30\S}$\href{http://orcid.org/0000-0002-5924-2464}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Marcel Stimberg}$^{31\ddagger}$\href{http://orcid.org/0000-0002-2648-4790}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Frank Stollmeier}$^{32\S}$\href{http://orcid.org/0000-0003-4858-0895}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Federico Vaggi}$^{33\ddagger}$\href{http://orcid.org/0000-0001-8100-158X}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Guillaume Viejo}$^{34\S}$,
%%
\textbf{Julien Vitay}$^{35\ddagger\S}$\href{http://orcid.org/0000-0001-5229-2349}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Anya Vostinar}$^{36\ddagger}$\href{http://orcid.org/0000-0001-7216-5283}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Roman Yurchak}$^{37\ddagger}$\href{http://orcid.org/0000-0002-2565-4444}{\includegraphics[width=8pt]{orcid}},
%%
\textbf{Tiziano Zito}$^{38\dagger}$\\