### 2017-07-10 affiliations.tex
Updated authors (because of new submission)

$^{1}$INRIA Bordeaux Sud-Ouest Talence, France – Institut des Maladies Neurodégénératives, Université de Bordeaux, CNRS UMR 5293, Bordeaux, France – LaBRI, Université de Bordeaux, Bordeaux INP, CNRS UMR 5800, Talence, France
%%
$^{2}$Centre de Biophysique Moléculaire, CNRS UPR4301, Orléans, France -- Synchrotron SOLEIL, Division Expériences, Gif sur Yvette, France
%%
$^{3}$Department of Electronic Systems, Faculty of Engineering and Science, Aalborg University, Denmark
%%
$^{4}$Department of Mechanical and Aerospace Engineering, the George Washington University, Washington DC, USA
%%
$^{5}$Department of Population Health and Reproduction, University of California Davis, Davis, CA, USA
%%
$^{6}$Instituut voor Theoretische Fysica, KU Leuven, Belgium -- PdB is a postdoctoral fellow of the Research Foundation -- Flanders (FWO)
%%
$^{7}$Laboratoire d'Informatique de l'Université du Maine (LIUM), Le Mans, France
%%
$^{8}$Unité de Neurosciences, Information et Complexité, Centre National de la Recherche Scientifique, Gif-sur-Yvette, France
%%
$^{9}$IGBMC, INSERM U964, CNRS UMR 7104, Université de Strasbourg, Illkirch, France
%%
$^{10}$Department of Cognitive Sciences, University of California Irvine, USA
%%
$^{11}$Computational Engineering and Design, Faculty of Engineering and the Environment, University of Southampton, UK
%%
$^{12}$Department of Computer Science, Humboldt-Universität zu Berlin
%%
$^{13}$Friedman Brain Institute, Icahn School of Medicine at Mount Sinai, NY, USA
%%
$^{14}$Institut des Systèmes Intelligent et de Robotique, Sorbonne Universités, UPMC Univ Paris 06, CNRS, Paris, France
%%
$^{15}$Experimental Psychology, University College London, UK
%%
$^{16}$UCL Great Ormond St Institute of Child Health. University College London, London, UK
%%
$^{17}$MRC Cognition and Brain Sciences Unit, Cambridge, UK
%%
$^{18}$Independent scholar, Enschede, The Netherlands
%%
$^{19}$Swammerdam Institute for Life Sciences, Center for Neuroscience, Faculty of Science, University of Amsterdam, Amsterdam, the Netherlands
%%
$^{20}$Department of Biology, College of Charleston, Charleston, SC, USA
%%
$^{21}$Department of Evolutionary Biology and Environmental Studies, University of Zurich, Switzerland
%%
$^{22}$Faculty of Science and Technology, Norwegian University of Life Sciences, Aas, Norway -- Institute of Neuroscience and Medicine (INM-6), Jülich Research Centre, Jülich, Germany
%%
$^{23}$Département de Sciences Biologiques, Université de Montréal, Montréal, Canada
%%
$^{24}$Berkeley Institute for Data Science, University of California Berkeley, Berkeley, CA, USA
%%
$^{25}$Department of Biology, Stanford University, Stanford, CA, USA
%%
$^{26}$Centre for Integrative Neuroscience, School of Psychology, University of Reading, UK
%%
$^{27}$Institute of Neurology, University College London, UK
%%
$^{28}$Institute of Neuroscience \& Medicine (INM-6) and Institute for Advanced Simulation (IAS-6) -- JARA-Brain Institute I (INM-10), Jülich Research Center, Jülich, Germany
%%
$^{29}$Department of Biology, University of Ottawa, Ottawa, Ontario, Canada
%%
$^{30}$Department of Fisheries and Wildlife, Michigan State University, MI, USA
%%
$^{31}$Sorbonne Universités, UPMC Univ Paris 06, INSERM, CNRS, Institut de la Vision, Paris, France
%%
$^{32}$Network Dynamics, Max Planck Institute for Dynamics and Self-Organization, Germany
%%
$^{33}$Amazon, Seattle, USA
%%
$^{34}$Montreal Neurological Institute, McGill University, Montreal, Canada
%%
$^{35}$Professorship for Artificial Intelligence, Department of Computer Science, Chemnitz University of Technology, Chemnitz, Germany
%%
$^{36}$Department of Computer Science, Grinnell College, IA, USA
%%
$^{37}$Symerio, Palaiseau, France
%%
$^{38}$Neural Information Processing Group, University of Tübingen, Germany
%%
\par