### 2017-09-28 affiliations-overleaf.tex
Added necessary script for overleaf/F1000 sublmission

\affil[1]{INRIA Bordeaux Sud-Ouest Talence, France – Institut des Maladies Neurodégénératives, Université de Bordeaux, CNRS UMR 5293, Bordeaux, France – LaBRI, Université de Bordeaux, Bordeaux INP, CNRS UMR 5800, Talence, France}
\affil[2]{Centre de Biophysique Moléculaire, CNRS UPR4301, Orléans, France -- Synchrotron SOLEIL, Division Expériences, Gif sur Yvette, France}
\affil[3]{Department of Electronic Systems, Faculty of Engineering and Science, Aalborg University, Denmark}
\affil[4]{Department of Mechanical and Aerospace Engineering, the George Washington University, Washington DC, USA}
\affil[5]{Department of Population Health and Reproduction, University of California Davis, Davis, CA, USA}
\affil[6]{Instituut voor Theoretische Fysica, KU Leuven, Belgium -- PdB is a postdoctoral fellow of the Research Foundation -- Flanders (FWO)}
\affil[7]{Laboratoire d'Informatique de l'Université du Maine (LIUM), Le Mans, France}
\affil[8]{Unité de Neurosciences, Information et Complexité, CNRS FRE 3693, Gif sur Yvette, France}
\affil[9]{Institut de Génétique et de Biologie Moléculaire et Cellulaire, INSERM U964, CNRS UMR 7104, Université de Strasbourg, Illkirch, France}
\affil[10]{Department of Cognitive Sciences, University of California Irvine, USA}
\affil[11]{Computational Engineering and Design, Faculty of Engineering and the Environment, University of Southampton, UK}
\affil[12]{Department of Computer Science, Humboldt-Universität zu Berlin}
\affil[13]{Friedman Brain Institute, Icahn School of Medicine at Mount Sinai, NY, USA}
\affil[14]{Institut des Systèmes Intelligent et de Robotique, Sorbonne Universités, UPMC Univ Paris 06, CNRS, Paris, France}
\affil[15]{Experimental Psychology, University College London, UK}
\affil[16]{UCL Great Ormond St Institute of Child Health. University College London, London, UK}
\affil[17]{MRC Cognition and Brain Sciences Unit, Cambridge, UK}
\affil[18]{Department of Ecology and Evolution, University of Lausanne, Lausanne, Switzerland}
\affil[19]{Independent scholar, Enschede, The Netherlands}
\affil[20]{BioMediTech Institute and Faculty of Biomedical Sciences and Engineering, Tampere University of Technology, Tampere, Finland}
\affil[21]{Swammerdam Institute for Life Sciences, Center for Neuroscience, Faculty of Science, University of Amsterdam, Amsterdam, the Netherlands}
\affil[22]{Department of Biology, College of Charleston, Charleston, SC, USA}
\affil[23]{Centre for Computer Science and Informatics Research, University of Hertfordshire, UK}
\affil[24]{Department of Evolutionary Biology and Environmental Studies, University of Zurich, Switzerland}
\affil[25]{Faculty of Science and Technology, Norwegian University of Life Sciences, Aas, Norway -- Institute of Neuroscience and Medicine (INM-6), Jülich Research Centre, Jülich, Germany}
\affil[26]{Département de Sciences Biologiques, Université de Montréal, Montréal, Canada}
\affil[27]{Berkeley Institute for Data Science, University of California Berkeley, Berkeley, CA, USA}
\affil[28]{Department of Biology, Stanford University, Stanford, CA, USA}
\affil[29]{Centre for Integrative Neuroscience, School of Psychology, University of Reading, UK}
\affil[30]{Institute of Neurology, University College London, UK}
\affil[31]{Institute of Neuroscience \& Medicine (INM-6) and Institute for Advanced Simulation (IAS-6) -- JARA-Brain Institute I (INM-10), Jülich Research Center, Jülich, Germany}
\affil[32]{Department of Biology, University of Ottawa, Ottawa, Ontario, Canada}
\affil[33]{Department of Fisheries and Wildlife, Michigan State University, MI, USA}
\affil[34]{Sorbonne Universités, UPMC Univ Paris 06, INSERM, CNRS, Institut de la Vision, Paris, France}
\affil[35]{Network Dynamics, Max Planck Institute for Dynamics and Self-Organization, Germany}
\affil[36]{Amazon, Seattle, USA}
\affil[37]{Montreal Neurological Institute, McGill University, Montreal, Canada}
\affil[38]{Professorship for Artificial Intelligence, Department of Computer Science, Chemnitz University of Technology, Chemnitz, Germany}
\affil[39]{Department of Computer Science, Grinnell College, IA, USA}
\affil[40]{Symerio, Palaiseau, France}
\affil[41]{Neural Information Processing Group, University of Tübingen, Germany}
\par