### 2017-11-07 replies_to_reviewers.md
Changes in reply to reviewer 2

# Replies to the reviewers' comments

## In reply to reviewer 1

## In reply to reviewer 2

We thank the reviewer for his enthusiastic comments, and we would be
happy to welcome him as a contributor to ReScience.

We have corrected the name of Jonathan Buckheit.