### 2017-11-08 replies_to_reviewers.md
Reproducibility of software environments
In reply to reviwer #1

# Replies to the reviewers' comments

## In reply to reviewer 1

We thank the reviewer for the in-depth examination of our work and for his useful suggestions for the future evolution of ReScience.

Concerning what the reviewer calls a defeatist attitude towards computational reproducibility, that is not at all how we see our approach. The manuscript focuses on reporting our experiences with ReScience. Concerning reproducibility, we note that having it verified by reviewers is not sufficient. We are not aware of any practical experience demonstrating that some other approach is efficient, which is why we consider the reproducibility of software environments an unsolved problem. Many ideas are being explored at the moment, including containerization as the reviewer mentions. The effectiveness of these ideas cannot be judged yet. Nobody knows, for example, if five years from now, the then-current version Docker will still be able to handle today's container images. We see our attitude not as defeatist, but as prudent. Many of the authors are contributing to the development and evaluation of new ideas, and we will certainly try some of them out in the context of ReScience.

We have reworded some of our statements to make clearer that our doubts about reproducibility apply only to today's state of the start. We have also added a new paragraph on the reproducibility of software environments to the Outlook section.

## In reply to reviewer 2

We thank the reviewer for his enthusiastic comments, and we would be
happy to welcome him as a contributor to ReScience.

We have corrected the name of Jonathan Buckheit.